---
title: "Problem Set 4: Key to Computational Problems"
author: | 
        | Amna Javed, Cory Briar, Sichao Jiang, Youssef Ait Benasser
        | EC 320: Introduction to Econometrics
        | University of Oregon
date: "Winter 2020"
output: 
  html_document:
    theme: flatly  
    highlight: tango
    includes:
      after_body: 
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F, cache = F)
```

## Preliminaries

```{r}
library(pacman)
p_load(stargazer, broom, tidyverse)

data <- read_csv("apple.csv")
```

## Exercise 1

*Run a regression of* `reglbs` *on* `regprc`*. Interpret the slope coefficient. Is the sign of the slope consistent with what you know about demand curves?*

```{r}
reg1 <- lm(reglbs ~ regprc, data = data)
tidy(reg1)
```

The slope estimate is -0.688, which tells us that the quantity of regular apples falls by 0.688 for each dollar-increase in the price of regular apples. In introductory microeconomics, you learned that demand curves are downward sloping. The negative slope estimate is consistent with the notion of downward-sloping demand.

## Exercise 2

*Run a regression of* `ecolbs` *on* `ecoprc`*. Interpret the intercept coefficient.*

```{r}
reg2 <- lm(ecolbs ~ ecoprc, data = data)
tidy(reg2)
```

The intercept estimate is 2.39, which tells us that consumers demand 2.39 pounds of eco-labeled apples when the price of eco-labeled apples is zero.

## Exercise 3

*Run a regression of* `reglbs` *on* `regprc` *and* `ecoprc`*. How does the estimated coefficient on* `regprc` *change? What does this tell you about the correlation between* `regprc` *and* `ecoprc`*? Justify your answer and then use* `R` *to verify.*

```{r}
reg3 <- lm(reglbs ~ regprc + ecoprc, data = data)
tidy(reg3)
```

The estimated slope falls from -0.688 to -1.57, which suggests that the omission of `ecoprc` in the first regression caused positive omitted-variable bias (you can show this with the first omitted-variable bias formula). The slope estimate on `ecoprc` is positive, so it must be the case that the correlation between `regprc` and `ecoprc` is positive (you can show this using the second omitted-variable bias formula). 

To verify the correlation, you can run a regression of `regprc` on `ecoprc` and check the sign of the slope estimate or simply calculate the correlation coefficient.

```{r}
# check the correlation (approach 1)
tidy(lm(regprc ~ ecoprc, data = data))

# check the correlation (approach 2)
cor(data$regprc, data$ecoprc)
```

As predicted by the omitted-variable bias formula, the correlation is positive.

## Exercise 4

*Run a regression of* `ecolbs` *on* `regprc` *and* `ecoprc`*. Identify and interpret* $R^2$*.*

```{r}
reg4 <- lm(ecolbs ~ regprc + ecoprc, data = data)
glance(reg4)
```

$R^2$ is 0.0364, which tells us that the various prices of apples explain 3.6 percent of the variation in the quantity demanded of regular apples.

## Exercise 5

*Summarize your regression results in a table. Which two of the four regressions have the highest* $R^2$*? Why?*

```{r, results='asis'}
stargazer(reg1, reg2, reg3, reg4, type = "html", keep.stat = c("rsq", "n"),
          notes = "<em>&#42;p&lt;0.1;&#42;&#42;p&lt;0.05;&#42;&#42;&#42;p&lt;0.01</em>",
          notes.append = FALSE)
```

<br>

**Note 1:** To ensure that R Markdown renders the table, you must specify `results='asis'` as a code chunk option.

**Note 2:** The default `stargazer` table includes stars (`*`, `**`, and `***`) in the table notes, but R Markdown mistakes these to mean *"italicize"* or *"bold"*. I include raw HTML in the `notes` argument to fix this bug. 

The regressions in columns 2 and 4 have the highest $R^2$. In this particular example, it appears that consumers are more price-sensitive to eco-labeled apples than they are to the prices of regular apples. In other words, prices explain more of the variation in `ecolbs` than `reglbs`, which likely explains why regressions 2 and 4 have the highest $R^2$. Between those two regressions, the regression in column 4 has a higher $R^2$ because there is an additional variable.

## Exercise 6

*Construct a 99 percent confidence interval for the* `ecoprc` *coefficient from the regression described in exercise 4.*

```{r}
tidy(reg4, conf.int = T, conf.level = 0.99)
```

The 99 percent confidence interval for the `ecoprc` coefficient is $-4.45 < \beta_\text{ecoprc} < -1.41$.